---
title: "Fertilizer Data Analysis"
output: html_document
date: "2026-04-29"
---

```{r setup, include=FALSE}
# Load necessary libraries
library(tidyverse)
library(knitr)
library(ggpubr)
```

## 1. Data Preparation
We are organizing the dataset into a "long format" where each row represents one sample.

```{r data}
yield_data <- data.frame(
  Fertilizer = factor(rep(c("X", "Y", "Z"), each = 20)),
  Crop = factor(rep(rep(c("Wheat", "Corn", "Soy", "Rice"), each = 5), 3)),
  Yield = c(
    # Blend X
    123, 156, 112, 100, 168, 128, 150, 174, 116, 109, 166, 178, 187, 153, 195, 151, 125, 117, 155, 158,
    # Blend Y
    135, 130, 176, 120, 155, 175, 132, 120, 187, 184, 140, 145, 159, 131, 126, 167, 183, 142, 167, 168,
    # Blend Z
    156, 180, 147, 146, 193, 186, 138, 178, 176, 190, 185, 206, 188, 165, 188, 175, 173, 154, 191, 169
  )
)
```

## 2. Descriptive Statistics
Below are the mean yields and standard deviations for each treatment group.

```{r descriptives}
descriptive_stats <- yield_data %>%
  group_by(Fertilizer, Crop) %>%
  summarise(
    Mean = mean(Yield),
    SD = sd(Yield),
    n = n(),
    .groups = 'drop'
  )

kable(descriptive_stats, caption = "Table 1: Descriptive Statistics by Group")
```

## 3. Two-Way ANOVA Results
We are testing the effect of Fertilizer, Crop, and their Interaction on Yield.

```{r anova}
# Run the ANOVA model
model <- aov(Yield ~ Fertilizer * Crop, data = yield_data)

# Display the ANOVA table
summary(model)
```

## 4. Interaction Plot
Visualizing how the performance of each fertilizer changes depending on the crop.

```{r interaction_plot, fig.width=8, fig.height=5}
ggline(yield_data, x = "Crop", y = "Yield", color = "Fertilizer",
       add = "mean_se",
       palette = "Set1",
       title = "Mean Yield by Fertilizer and Crop",
       xlab = "Crop Type", ylab = "Mean Yield (+/- SE)")
```
